\documentclass[./dissertation.tex]{subfiles}
\begin{document}

  \chapter*{Acknowledgements}

  %This will add the acknowledgements back to the toc
  \addcontentsline{toc}{chapter}{Acknowledgements}
I would like to express my deepest gratitude and appreciation to the following individuals, without whom the successful completion of my PhD thesis would not have been possible.

Firstly, I am immensely grateful to Pierangela Bardazzi Mattei, for her timely intervention on the day of my interview, waking me up and ensuring that I caught the train to this life-changing opportunity.

I would like to sincerely thank my academic supervisor, Prof. Giorgio di Natale, at the University of Grenoble Alpes, for his unwavering guidance, support, and availability throughout my doctoral journey.

My heartfelt appreciation goes to Jean Marc Daveau for his invaluable mentorship in the industrial environment of ST Microelectronics, providing me with essential insights and expertise.

I am deeply grateful to Philippe Roche for his adept management and for sharing his profound understanding of what it takes to run a department from a managerial perspective.

I would like to extend my gratitude to APSYS AIRBUS, particularly to Emmanuel Arbaretier, Marc Maitre, and Thomas Jaquet, for their indispensable technical support, which was instrumental in the successful recomposition of data.

My appreciation also goes to the CERN staff, including Kostas Klukinas, Aditha Pully, Stefano Esposito, Roberto Beccherle, and others, for providing me with the opportunity to work in such a stimulating environment.

I am thankful for the camaraderie and assistance provided by my colleagues at ST Microelectronics, who have contributed both technically and personally to my development.

Lastly, I would like to express my gratitude to my colleagues at UGA, particularly Pietro Inglese and Sergio Vinagrero Gutierrez, for their unwavering support, friendship, and technical expertise during my PhD journey.

Once again, I extend my deepest thanks to all who have supported me throughout this incredible endeavor.

\end{document}
