\documentclass[./dissertation.tex]{subfiles}
\begin{document}


\chapter{Failure Mode and Effect Analysis FMEA}

\section{Introduction}
\subsection{General}
FMEA and FMECA are important techniques for a reliability assurance programme.They can be applied to a wide range of problems which may occur in technical systems, and can be carried out in varying degrees of depth, or modified, to suit a particular purpose. The analysis is carried out in a limited way during the conception, planning, and definition phases and more fully in the design and development phase. It is however important to remember that the FMEA is only part of a reliability and
maintainability programme which requires many different tasks and activities. FMEA is an inductive method of performing a qualitative system reliability or safety analysis from a low to a high level. A thorough understanding of the system under analysis is essential prior to undertaking FMEA. Functional diagrams and other system drawings are normally necessary for this understanding. Reliability block diagrams, fault trees and/or state diagrams are then usually derived from these in order to
carry out the analysis. In many instances the block diagram descriptions and block diagram failure descriptions are included in the FMEA format. Separate diagrarns will be needed for the
following:
\begin{enumerate}
    \item The way in which different criteria for system faiulre are determined;
    \item Degradation of function or reduction in assurance of function;
    \item Alternative operational phases
\end{enumerate}

\subsection{Purpose of the Analysis}
The reasons for undertaking FMEA (or FMECA) may include the following:

\begin{itemize}
    \item to identify those failures which have unwanted effects on system operation, e.g. safety critical failures;
    \item to satisfy contractual conditions that an FMEA should be completed;
    \item where appropriate, to quantify the reliability and/or safety of the system;
    \item to allow improvements of the system's reliability and/or safety (e.g. by design or quality assurance action)
    \item to produce aids to fault diagnosis;
    \item to allow improvement of the system's maintâinability (by highlighting areas of risk or non-conformance for maintainability).
\end{itemize}

ln view of these reasons the objectives of an FMEA (or FMECA) may include the following: 

\begin{enumerate}
    \item a comprehensive identification and evaluation of all the unwanted effects within the defined boundaries of the system being analysed, and the sequences of events brought about by each identified item failure mode, from whatever cause, at various levels of the system's functional hierarchy;
    \item the determination of the significance (or criticality) of each failure mode with respect to the system's correct function or performance and the impact on the reliability and/or safety of the process concerned;
    \item a classification of identified failure modes according to relevant characteristics, including detectability, diagnosability, testability, item replaceability, compensating and operating provisions (repair, maintenance, logistics, etp.);
    \item an estimation of measures of the significance and probability of failure.
\end{enumerate}


\subsection{Basic Principles of FMEA}
The following concepts are essential to FMEA:

\begin{enumerate}
    \item breakdown of the system into 'elements';
    \item a diagram of the system's functional structure and identification of the various data which are needed to perform the FMEA;
    \item the failure mode concept (a part may have several failure modes or a failure mode may involve several parts);
    \item identification of new physical features or new requirements;
    \item the criticality concept and the measure to be used (if criticality analysis is required).
\end{enumerate}

Further, it is essential to specify the existing links between the FMEA (and the FMECA) and other qualitative (and quantitative) analytical methods within the overall reliability programme.  Very few designs are wholly new. Most are to some extent developments of old designs. FMEA should use the information on existing systems and draw attention to the need for tests, etc. for the new parts.  



\section{Procedure}

\subsection{General}
The wide variation in complexity of system designs and applications may require the development of highly individualized FMEA procedures consistent with the information available. Traditionally, there have been wide variations in the manner in which FMEA is conducted and presented. However, the analysis is usually done in a standard manner and presented on a worksheet that contains a core of essential information which can be developed and extended to suit the particular system or project to which it is applied. A typical example of a worksheet is shown in Figure 1.

The procedure consists of the following four main stages:

\begin{enumerate}
\item Preparatory definition of the system including the design, functional, operational, maintenance, and environmental requirements;
\item Establishment of the basic principles and purposes of the FMEA and the form of its presentation;
\item Carrying out the FMEA using the appropriate worksheet designed according to (a) and (b);
\item Reporting of the complete analysis including any conclusions and recommendations made.
\end{enumerate}

A more detailed consideration of the information needed is given in Section 2.4.2.

\subsection{Preparation}
At the commencement of an analysis, the following preparations should be made:
\begin{enumerate}
\item The analyst should have available the information listed in Section 2.4.2.2 to 2.4.2.7 that clearly defines the system to be analyzed.
\item It will usually be necessary for the analyst to translate the information into some form of functional, hierarchical, or reliability block diagrams. An example of a functional diagram is shown in Figure 2. This diagram shows how the failure effects at the part level form the failure modes at the module level, the failure effects at the module level form the failure modes at the subsystem level, and so on. Such a representation of the system should explicitly identify the system's functional structure, the system boundary, and the inputs and outputs crossing that boundary. Further information is given in Section 2.4.2.8 to 2.4.2.10.
\end{enumerate}

\subsection{FMEA principles}
The following principles should be applied:
\begin{enumerate}
\item Define clearly the purposes and uses of the FMEA as indicated in Section 2.1.2.
\item Establish and define the relationships with other forms of reliability analysis with which the FMEA may subsequently be integrated. (See Section 2.3.5.)
\item Define the scope of the FMEA in relation to the functional structure and hierarchical structure of the system as described by the block diagrams referred to in Section 2.4.2.10. It is essential to define the lowest level in the system's hierarchical structure at which the analysis will start. The guidance given in Sections 2.3.4, 2.4.1, and 2.4.2.8 is especially important for this task.
\item Define the format of the FMEA worksheet to suit the project requirements. The core information considered essential is as follows:
\end{enumerate}

\begin{enumerate}
\item The name of the item in the system being analyzed;
\item Function performed by the item;
\item Identification number of the item;
\item Failure modes of the item;
\item Failure causes;
\item Failure effects on the system;
\item Failure detection methods;
\item Compensating provisions;
\item Severity of effects;
\item Remarks.
\end{enumerate}
Other information required for the particular system and project needs to be defined by the analyst according to the purposes of the


\end{document}
