\documentclass[./dissertation.tex]{subfiles}
\begin{document}
    \chapter{Conclusion}
In conclusion, this PhD thesis may have not made significant and practical contributions to the field, but by addressing critical gaps in knowledge, developing innovative methodologies, and providing valuable insights paved the way for further research. The Field of Automatic Certification and safety assessment is rather open field, joinin the specialized conferences and symposiums it is clear that the community it is, yes willing, but not read to yield the responsability of such and analyis to an automated method. Nevertheless the rigorous theoretical framework presented in the thesis, which integrates and builds upon previous literature, serves as a solid foundation for the empirical analysis future PhD will continue, under different spotlights to enrich this research, until finally it will be the state of the art for the relibility assessment of electronic components. In any case, the extensive data collection, meticulous experimental design, and robust validation of the proposed models have allowed me to have a more comprehensive understanding of the subject matter, thus enhancing the existing body of knowledge and allowing a completlely different vision of the field.

%%%%%%%%%%%%%%%%%%%%%%%%%%% NEW PART

The research conducted in this thesis has successfully adapted and extended the fault tree-based extraction of failure metrics, which has conventionally been applied to macroscopic electromechanical systems, to the realm of automotive SoCs. The current state-of-the-art in automotive SoCs involves largely manual metric extraction and expert dependence, whereas the verification of countermeasure effectiveness is typically performed through targeted fault injection on specific sub-parts of the system or irradiation under a particle beam.

The first part of the thesis involved an in-depth literature review of fault tree construction, the ISO26262 standard, and the derivation of different reliability metrics for digital SoC-type systems. The developed methodology extracts metrics at the block level using two different methods: an analytical method based on probabilities and an experimental method based on fault injection. The focus of this research was not to create new probability codes or fault injection tools, but rather to establish a methodology for employing these techniques in the context of an SoC to obtain the desired data.

The second part of the thesis addressed the composition of data obtained at the functional block level to derive the ISO26262 metrics at the system level (SoC). The developed composition method was tailored to the unique characteristics of SoCs, such as their communication systems, real-time calculations, and fault models imposed by the ISO26262 standard.

In the third part of the thesis, the developed methodology was applied to an SoC-type system, and the results were validated. This practical application of the research demonstrates the effectiveness of the proposed methodology in a real-world context, offering valuable insights and facilitating the ISO26262 certification process for automotive SoCs.

Through this research journey, the thesis has contributed to bridging the gap between traditional fault tree-based methods and their application to the automotive SoC domain. The findings have not only enhanced the existing literature but also provided practical implications for stakeholders, policymakers, and practitioners involved in the development and certification of automotive SoCs.

As with any research endeavor, this thesis has also identified areas for future exploration and improvement. Future researchers can build upon the foundations laid in this work to further refine the proposed methodology, expand its applicability to other domains, and develop innovative techniques for extracting and analyzing reliability metrics in complex systems.

In conclusion, this thesis marks a significant milestone in the field of automotive SoC reliability and ISO26262 certification. The rigorous theoretical framework, extensive empirical analysis, and robust validation of the proposed models have provided a comprehensive understanding of the subject matter, paving the way for further research and advancements. The interdisciplinary nature of the study has fostered collaborations and synergies among different domains, encouraging a more holistic approach to problem-solving. As the field continues to evolve, the knowledge generated by this research will serve as a stepping stone for further innovations and the development of more efficient and reliable automotive systems.





%%%%%%%%%%%%%%%%%%%%%%%% MAYBE
Throughout this research journey, several challenges were encountered and overcome, leading to a more profound appreciation for the complexities inherent in the chosen field. The interdisciplinary nature of the study has fostered collaborations and synergies among different domains, encouraging a more holistic approach to problem-solving. By embracing a multidisciplinary perspective, this thesis has successfully transcended the traditional boundaries of the field and provided innovative solutions to complex issues.

The findings of this research have also offered practical implications for stakeholders, policymakers, and practitioners. The policy recommendations and best practices derived from the results can be applied to various real-world contexts, contributing to more effective decision-making and the development of more sustainable and equitable systems. Moreover, the thesis has highlighted the importance of continuous dialogue between academia and industry, emphasizing the need for a collaborative approach to address the multifaceted challenges faced by our society.

In addition to the immediate contributions, this thesis has also identified several areas for future research. The findings presented here open up new avenues for exploration, inviting future researchers to delve deeper into the subject matter and build upon the foundations laid in this work. As the field continues to evolve, it is expected that the knowledge generated by this research will serve as a stepping stone for further innovations and advancements, leading to the discovery of new solutions and a better understanding of the complexities of our world.

Furthermore, the methodological advancements and novel approaches proposed in this thesis offer exciting opportunities for researchers in other domains. The cross-disciplinary applicability of these methods can be leveraged to tackle challenges in diverse areas, ultimately contributing to the growth of knowledge across fields.

Throughout the research process, the importance of ethical considerations and the impact of the study on society have been acknowledged and addressed. This thesis has demonstrated that the pursuit of knowledge must be accompanied by a commitment to responsible research practices, ensuring that the generated knowledge contributes positively to the well-being of all stakeholders involved.


This thesis is a testament to the power of collaboration, perseverance, and the relentless pursuit of knowledge in the service of a better understanding of our world. As I embark on the next chapter of my academic journey, I carry with me the lessons learned, the skills acquired, and the passion for discovery that has been nurtured throughout the course of my doctoral studies. I am committed to using my research as a catalyst for positive change and to continuing my lifelong pursuit of knowledge in the service of society.
\end{document}
